Thesis Proposal

### 2019-10-24 ResearchReport/article.tex
rename folders
Added a simulation folder too

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
@

@@

% For a simple illustration of basic Poisson and NB count regression the
% \code{quine} data from the \pkg{MASS} package is used. This provides the number
% of \code{Days} that children were absent from school in Australia in a
% particular year, along with several covariates that can be employed as regressors.
% The data can be loaded by
%
% <<data>>=
% data("quine", package = "MASS")
@
%
% and a basic frequency distribution of the response variable is displayed in
% Figure~\ref{fig:quine}.

@@

% \begin{figure}[t!]
% \centering
% <<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
% par(mar = c(4, 4, 1, 1))
% plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
% axis(2)
% axis(1, at = 0:16 * 5, labels = FALSE)
% axis(1, at = 0:8 * 10)
% box()
@
% \caption{\label{fig:quine} Frequency distribution for number of days absent
% from school.}
% \end{figure}

% As a first model for the \code{quine} data, we fit the basic Poisson regression
% model. (Note that JSS prefers when the second line of code is indented by two
% spaces.)
%
% <<poisson>>=
% m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
%   family = poisson)
@
%
% To account for potential overdispersion we also consider a negative binomial
% GLM.
%
% <<negbin>>=
% library("MASS")
% m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
@
%
% In a comparison with the BIC the latter model is clearly preferred.
% %
% <<comparison>>=
% BIC(m_pois, m_nbin)
@
% %
% Hence, the full summary of that model is shown below.
% %
% <<summary>>=
% summary(m_nbin)
@

@@

The results in this paper were obtained using \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the \pkg{mice}~\Sexpr{packageVersion("mice")} package. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.

>>>>

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}

@@

% For a simple illustration of basic Poisson and NB count regression the
% \code{quine} data from the \pkg{MASS} package is used. This provides the number
% of \code{Days} that children were absent from school in Australia in a
% particular year, along with several covariates that can be employed as regressors.
% The data can be loaded by
%
% <<data>>=
% data("quine", package = "MASS")
%
% and a basic frequency distribution of the response variable is displayed in
% Figure~\ref{fig:quine}.

@@

% \begin{figure}[t!]
% \centering
% <<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
% par(mar = c(4, 4, 1, 1))
% plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
% axis(2)
% axis(1, at = 0:16 * 5, labels = FALSE)
% axis(1, at = 0:8 * 10)
% box()
% \caption{\label{fig:quine} Frequency distribution for number of days absent
% from school.}
% \end{figure}

% As a first model for the \code{quine} data, we fit the basic Poisson regression
% model. (Note that JSS prefers when the second line of code is indented by two
% spaces.)
%
% <<poisson>>=
% m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
%   family = poisson)
%
% To account for potential overdispersion we also consider a negative binomial
% GLM.
%
% <<negbin>>=
% library("MASS")
% m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
%
% In a comparison with the BIC the latter model is clearly preferred.
% %
% <<comparison>>=
% BIC(m_pois, m_nbin)
% %
% Hence, the full summary of that model is shown below.
% %
% <<summary>>=
% summary(m_nbin)

@@

The results in this paper were obtained using \proglang{R}~3.6.1 with the \pkg{mice}~3.6.0.9000 package. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.

### 2019-10-24 ThesisProposal/ThesisProposal.Rmd
rename folders
Added a simulation folder too

Word count: 713

@@

With MI, many assumptions are made about the nature of the observed and missing parts of the data and their relation to the 'true' data generating model [@buur18]. Without proper evaluation of the imputations and the underlying assumptions, any drawn inference may erroneously be deemed valid. Such evaluation measures are currently missing or under-developed in MI software, like the world leading `R` package `MICE` [@mice11]. Therefore, I will answer the following question: 'Which measures are vital for evaluating the validity of multiply imputed data?'.

@@

The validity of the MI solution depends on numerous assumptions that cannot be verified from the observed data alone. Consequently, existing evaluation methods rely on **[rephrase: proxy measures]**. For the following assumptions, no reliable *proxy measures* have been proposed and/or implemented: 1) *ignorability* of the *missingness mechanism*; 2) congruity of the *imputation models*; and 3) *compatibility* of the MI modeling procedure. 
<!-- Rubin (1987b, 160-66) subdivided the work needed to create imputations into three tasks. The modeling task chooses a specific model for the data, the estimation task formulates the posterior parameters distribution given the model and the imputation task takes a random draws for the missing data by drawing successively from parameter and data distributions @buur18 par 4.5.1 -->

@@

1. A missingness mechanism is said to be ignorable when the sprobability to be missing does not depend on the missing data itself [@rubin87]. Evaluation of this assumption is usually done with sensitivity analyses to assess the robustness of inferences to violations. Some practical guidelines exist (e.g., [@nguy17]), but current MI software does not facilitate this methodology for empirical researchers.

@@

2. Congruent imputation models capture all required relations between observed and missing parts of the data. The exent to which this has been successful can be evaluated by plotting conditional distributions [@abay08]. Such visualizations are available in `MICE`, but subsequent statistical tests to quantify the relations with covariates are not provided. Additionally, there is potential to assess model fit by means of *over-imputation* [@buur18] or *double robustness* [@bang05].

@@

@@

Subsequently, I will use `R Shiny` [@shiny17] to implement the convergence indicator and other evaluation measures in `ShinyMICE`, see Figure 1. The application will at least contain methodology for: sensitivity analyses; data visualizations (e.g., scatter-plots, densities, cross-tabulations); and statistical evaluations of relations between variables pre- versus post-imputation (i.e., $\chi^2$-tests or $t$-tests).

>>>>

Word count: 709 words

@@

With MI, many assumptions are made about the nature of the observed and missing parts of the data and their relation to the 'true' *data generating model* [@buur18]. Without proper evaluation of the imputations and the underlying assumptions, any drawn inference may erroneously be deemed valid. Such evaluation measures are currently missing or under-developed in MI software, like the world leading `R` package `MICE` [@mice11]. Therefore, I will answer the following question: 'Which measures are vital for evaluating the validity of multiply imputed data?'.

@@

The validity of the MI solution depends on numerous assumptions that cannot be verified from the observed data alone. So instead of statistical tests for assumptions, evaluation procedures have been developed. For the following assumptions, no reliable procedure has been proposed and/or implemented: 1) *ignorability* of the *missingness mechanism* [@rubin87]; 2) *congeneality* of the imputation models [@meng94]; and 3) *compatibility* of the MI modeling procedure [@rubin96]. 
<!-- Rubin (1987b, 160-66) subdivided the work needed to create imputations into three tasks. The modeling task chooses a specific model for the data, the estimation task formulates the posterior parameters distribution given the model and the imputation task takes a random draws for the missing data by drawing successively from parameter and data distributions @buur18 par 4.5.1 -->

@@

1. A missingness mechanism is said to be ignorable when the probability to be missing does not depend on the missing data itself. Violation of this assumption can gravely affect inferences. Robustness of inferences to varying degrees of violation can be assessed with sensitivity analyses. Some practical guidelines exist (e.g., [@nguy17]), but current MI software does not facilitate this methodology for empirical researchers.

@@

2. Congeneal imputation models capture all required relations between observed and missing parts of the data. The extent to which this has been successful can be evaluated by plotting conditional distributions [@abay08]. Such visualizations are available in `MICE`, but subsequent statistical tests to quantify the relations with covariates are not provided. ^[Additionally, there is potential to assess model fit by means of *over-imputation* [@buur18] or *double robustness* [@bang05]-- topics that are only persued if time permits.]

@@

&nbsp;

@@

Subsequently, I will use `R Shiny` [@shiny17] to implement the convergence indicator and existing evaluation measures in `ShinyMICE`, see Figure 1. The application will at least contain methodology for: sensitivity analyses; data visualizations (e.g., scatter-plots, densities, cross-tabulations); and statistical evaluation of relations between variables pre- versus post-imputation (i.e., $\chi^2$-tests or $t$-tests).
